### 2021-10-04 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@1fb29d938b9c265612a7294ca72ccb9420809de7 🚀

In order to simplify the calculations, I make the assumptions that the
networks have equal species richness (noted \(S\)), so that
\(S_1 = S_2 = S\), and the same connectance (noted \(\rho\)), so that
\(\rho_1 = \rho_2 = \rho\). As a consequence, the two networks have the
same number of links \(L = \rho\times S_1 = \rho\times S_2\). The
assumption of equal connectance will be relaxed in a subsequent
numerical experiment. These simplifications allow to express the size of
\(C\), \(R\), and \(T\) only as functions of \(p\) and \(q\), as they
would all be multiplied by \(L\), which can therefore be dropped from
the calculation.

\begin{figure}
\hypertarget{fig:numexp1}{%
\centering
\includegraphics{figures/sharing_v_rewiring/components.png}
\caption{Values of \(\beta_{os}\), \(\beta_{wn}\), \(\beta_{st}\), and
\(\beta_{st}/\beta_{wn}\) as a function of the probability \(q\) or
sharing a link (\(x\)-axis), and the probability \(p\) of sharing a
species (\(y\)-axis). Larger values indicate \emph{more} dissimilarity,
such that for \(p=q=1\) the dissimilarity as measured by
\(\beta_{wn}=0\), and for \(p=q=0\) the dissimilarity as measured by
\(\beta_{wn}=1\).}\label{fig:numexp1}
}
\end{figure}

The rewiring component \(\beta_{os}\) varies as a function of the
proportion of shared links that are rewired; by contrast, \(\beta_{wn}\)
varies \emph{only} as a function of the proportion of links that are
shared: that the unshared links are established between common or unique
species has no effect on overall network dissimilarity. The quadratic
nature of the denominator for \(\beta_{st}\) is clear here, with a
maximum reach when there is no re-wiring, and a small number of shared
links (\emph{i.e.} the networks are almost entirely dissimilar except
for the links between shared species). Althought the \emph{raw} values
of \(\beta_{st}\) may seem low, the normalization using
\(\beta_{st}/\beta_{wn}\) magnifies this effect: its values are indeed
maximized when the rewiring is lower, \emph{i.e.} all of the network
variation stems from turnover processes.

@@

Consider two bipartite networks, each with \(R\) species on either side,
and each with the same connectance \(\rho\). We will assume that these
networks \emph{share} a proportion \(p\) of their species from one side
(and share all species from the other), and that the interactions
between these species are undergo rewiring with at a rate \(q\). This is
sufficient information to calculate the values of \(A\), \(S\), and
\(U\) required to get the values of \(\beta_{os}\) and \(\beta_{wn}\).
Note that the simplification of assuming that only species from one side
can vary is merely for the sake of simplicity, but does not decrease the
generality of the argument.

Each network will have \(\rho(1-p)R^2\) interactions that are unique due
to species turnover, and so

>>>>

In order to simplify the calculations, I make the assumptions that the
networks have equal species richness (noted \(S\)), so that
\(S_1 = S_2 = S\), and the same connectance (noted \(\rho\)), so that
\(\rho_1 = \rho_2 = \rho\). As a consequence, the two networks have the
same number of links \(L = \rho\times S_1^2 = \rho\times S_2^2\). The
assumption of equal connectance will be relaxed in a subsequent
numerical experiment. These simplifications allow to express the size of
\(C\), \(R\), and \(T\) only as functions of \(p\) and \(q\), as they
would all be multiplied by \(L\), which can therefore be dropped from
the calculation.

\begin{figure}
\hypertarget{fig:turnrew}{%
\centering
\includegraphics{figures/sharing_v_rewiring/components.png}
\caption{Values of \(\beta_{os}\), \(\beta_{wn}\), \(\beta_{st}\), and
\(\beta_{st}/\beta_{wn}\) as a function of the probability \(q\) or
sharing a link (\(x\)-axis), and the probability \(p\) of sharing a
species (\(y\)-axis). Larger values indicate \emph{more} dissimilarity,
such that for \(p=q=1\) the dissimilarity as measured by
\(\beta_{wn}=0\), and for \(p=q=0\) the dissimilarity as measured by
\(\beta_{wn}=1\). As expected, the relative importance of turnover
(\(\beta_{st}\)) is maximal when there is no rewiring, and when turnover
increases.}\label{fig:turnrew}
}
\end{figure}

The value of \(C\) is the proportion of shared species \(p^2\), as per
fig.~\ref{fig:conceptual}, times the proportion of shared links, \(q\),
giving \(C = qp^2\). Each network has \(r = p^2-(qp^2)\) rewired links,
which leads to \(R = 2r = 2p^2(1-q)\). Finally, we can get the number of
unique links in each network \(t\) by substracting \(C+r\) from the
total number of links (which, since we scale everything by \(L\), is 1),
yielding \(t = 1 - qp^2 - p^2 + qp^2\), which is \(t = 1-p^2\). The
total number of unique links due to turnover is \(T = 2t = 2(1-p^2)\).
It is important to note that \(C\) and \(R\), namely the number of links
that are kept or rewired, depends on species sharing (\(p\)), as the
possible size of the overlap between the two networks does, but the
quantity of links that are different due to turnover does not depends on
rewiring.

With the values of \(C\), \(R\), and \(T\), we can write

\[\beta_{os} = \frac{2p^2(1-q)}{2p^2q+2p^2(1-q)} = \frac{1-q}{q + 1 -q} =
(1-q)\,.\]

This is a first noteworthy result: the value of \(\beta_{os}\), in the
ideal scenario of equal links and richness, is the probability of link
re-wiring. Because this is true regardless of the value of \(p\)
(species turnover), this makes \(\beta_{os}\) a strongly ecologically
informative component.

Similarly, we can write

\[\beta_{wn} = \frac{2p^2(1-q)+2(1-p^2)}{2p^2q + 2p^2(1-q)+2(1-p^2)} = \frac{p^2(1-q)+(1-p^2)}{p^2q+p^2(1-q)+(1-p^2)} = 1-qp^2\,.\]

The overall dissimilarity responds to \(q\) (rewiring) linerarly, and to
\(p\) quadratically (which is expected assuming unipartite networks, in
which species are present on both sides).

Expressing \(\beta_{os}\) and \(\beta_{wn}\) as functions of \(p\) and
\(q\) trivializes the search for the expression of \(\beta_{st}\), which
is

\[\beta_{st} = 1 - p^2q - 1 + q = q\times(1-p^2)\,.\]

It is worth examining this solution in some detail. \(\beta_{st}\)
scales linearly with the probability that a link will \emph{not} be
rewired -- in other words, in a pair of networks for which rewiring is
important (\(q\) goes to 0), species turnover is going to be a
\emph{relatively} less important mechanism to dissimilarity.
\(\beta_{st}\) increases when turnover is important (\(p\) goes to 0),
and therefore \(\beta_{st}\) represents a \emph{balance} between species
turnover and link rewiring. These three values, as well as
\(\beta_{st}/\beta_{wn}\), are represented in fig.~\ref{fig:turnrew}.

@@

The results presented in fig.~\ref{fig:turnrew} include the strong
assumption that the two networks have equal connectance. Although the
range of connectances in nature tends to be very strongly conserved
within a system, we can relax this assumption, by letting one network
have more interactions than the other. Note that for the sake of
notation simplicity, I maintain the constraint that the two networks are
equally species rich. Therefore, the sole variation in this numerical
experiment is that one network has \(L_1 = \rho\times a\times S^2\), and
the other network has \(L_2 = \rho\times S^2\); in other words,
\(L_1 = a\times L\) and \(L_2 = L\). As one step of the components
calculations involves a \(\text{min}\) operation, I will add the
constraint that \(L_1 \le L_2\), which is to say \(0 < a \le 1\). The
value of \(a\) is the \emph{ratio} of connectances of the two networks,
and the terms \(S^2\) and \(\rho\) being shared across all factors, they
will be dropped from the calculations.

The maximal number of links that can be shared is \(ap^2\) (\emph{i.e.}
\(\text{min}(p^2, ap^2)\)), as we cannot share more links than are in
the sparsest of the two networks. Of these, \(q\) are not rewired,
leading to \(C = aqp^2\). The number of links that are rewired in
network 1 is the number of its links between shared species minus \(C\),
\emph{i.e.} \(r_1 = ap^2 - aqp^2 = ap^2(1-q)\), and similarly
\(r_2 = p^2 - aqp^2 = p^2(1-aq)\), leading to
\(R = r_1 + r_2 = p^2 \left[a(1-q)+1\right]\). Using the same approach,
we can get \(t_1 = a(1-p^2)\) and \(t_2 = (1-p^2)\), leading to
\(T = t_1 + t_2 = (1-p^2)(1+a)\).
